---
title: "Untitled"
author: "Your Name"
date: "September 28, 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
# Reads in the Harry Potter data set
hp <- read.csv("http://cddesja.github.io/classes/e411prma2015-3/data/hp.csv", header = T)[,-7]
```

# R Assignment 3

For this assignment, you'll be using item response theory. As with before, we'll use an example to create all the `R` code you'll need for this assignment. The data set used in this example is called `Scored` from the `irtoys` package and the data set you will use for the actual assignment is `hp`. __Please make sure you are online when you are doing this assignment otherwise you will get an error that says data set hp is not found__!

For this assignment, you need to install the `irtoys` package. To do this, run the following command once in RStudio:

install.packages("irtoys")

If you are on a Windows machine, please launch RStudio with Administrator privileges by right-clicking the RStudio icon, clicking "run as Administator" and then run the following commanded in RStudio:

install.packages("irtoys")

```{r, message =  FALSE, echo = FALSE}
library("irtoys")
```

__You will get an error if you haven't installed irtoys yet. So please make sure you do this before you proceed any further.__

```{r, eval = FALSE, echo = FALSE}

# Load the irtoys library
library("irtoys")

# Please note that if you run "Knit to HTML" you do not need to run that command below.
# I have already automatically included that in a Chunk for you.
# But if you are run R interactively (i.e. running each line of code prior to running the Knit command)
# Then you do need to run the library("irtoys") command

# Make sure to changed Scored for hp below !

#
# For Questions 1 - 5
#

# This runs the Rasch model.
rasch_model <- est(Scored, model="1PL", engine="ltm", rasch = TRUE)

#
# Estimated Parameters
rasch_params <- rasch_model$est
colnames(rasch_params) <- c("Discrimination", "Difficulty", "Guessing")
rownames(rasch_params) <- paste("Item", 1:ncol(Scored))
rasch_params

#
# Standard Errors
# 
est_se <- rasch_model$se
colnames(est_se) <- c("Discrimination SE", "Difficulty SE", "Guessing SE")
rownames(est_se) <- paste("Item", 1:ncol(Scored))
est_se

#
# Plot item response function for items 1, 3, 5
#
plot(irf(rasch_params[c(1,3,5),]), co = NA)  
# you can safely delete co = NA if you want the lines to all be black

#
# Estimated abilities
#
rasch_abl <- as.data.frame(mlebme(Scored, ip = rasch_params))

# This prints out the ability estimates
# est - is the estimated ability
# sem - is the standard error of measurement
# n - is the number of items that each person responded to
rasch_abl


min(rasch_abl$est)  # Prints the minimum score
max(rasch_abl$est)  # Prints the maximum scores
which.min(rasch_abl$est)  # Prints out the person with the minimum score
which.max(rasch_abl$est)  # Prints out the person id with the maximum score

#
# For Questions 6 - 7
#

# This code runs the 2-PL model
twopl <- est(Scored, model="2PL", engine="ltm")
twopl_params <- twopl$est
colnames(twopl_params) <- c("Discrimination", "Difficulty", "Guessing")
rownames(twopl_params) <- paste("Item", 1:ncol(Scored))
twopl_params

# Plot the IRFs for the 2-PL model.
plot(irf(twopl_params[c(1,3,5),]), co = NA)  

#
# Estimated abilities
#
twopl_abl <- as.data.frame(mlebme(Scored, ip = twopl_params))

# Correlation between the ability estimates
cor(twopl_abl$est,rasch_abl$est)
```

Please replace `Scored` with `hp` to answer the following questions.

## Questions

### Questions (1 - 4) are asking about the Rasch Model

#### Question 1. Which item was the easiest item and which item was the hardest? Which item was the second hardest? Do the results surprise you? Why? (3 pt) 

#### Question 2. What would we expect the probability of a correct response would be for someone who had an ability score of 0 for the easiest and the second hardest items? Hint: It might be useful to create an item reponse function plot to answer this question.  (2 pts)

#### Question 3.  What was the range in estimated abilities scores on the HP pretest?  (2 pts)

#### Question 4. Provide a 95% confidence interval for the estimated ability for the student who did the best on the test and interpret it. (2 pts)

### Questions 5 - 7 are asking you about the 2-PL 

#### Question 5. Which items would you decide to rewrite or remove from the pretest based on the 2-PL? Why would you rewrite or remove them? (2 pts)

#### Question 6. Provide a plot of items 8 and 9. Describe the differences in these two items' item response functions (3 pt).

#### Question 7  What is the correlation between the ability estimates on the Rasch model and the 2-PL? If your interest was solely on estimating person abilities, do you think you would draw the same conclusions from both models? Why? (2 pts)