---
title: 'R Assignment #1'
author: "Your Name"
date: "3 September 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1 - Describe the cars dataset

#### Question 1a: How many variables are in the cars dataset? (1 pt)
There are two variables in the cars dataset
```{r}
names(cars)
```

#### Question 1b: What are the names of the variables in the cars dataset and what do they mean? (2 pts)

The names of the variables in the cars data set are `speed` and `dist`. Speed is the speed in miles per hour and dist is the stopping distance in feet. This information was obtained from the help page

```{r}
?cars
names(cars)
```

#### Question 1c: How many observations are there in the cars dataset? (1 pt)
There are 50 observations in the cars dataset.
```{r}
nrow(cars)
```

#### Question 1d: Examine the descriptive statistics for the cars dataset. Please report the median and the mean for the variables and comment on what this implies about skewness. (3 pts)
The mean and median for `speed` are 15.4 and 15.0, respectively. Given that the mean is slightly above the median you might expect that the data are slightly postively (or right) skewed. However, the difference is not very large, so the data may be roughly symmetric.

The mean and median for `dist` are 42.98 and 36.00, respectively. Given that mean is a bit above the median, you might expect that the data are positively (or right) skewed. You would expect dist to be more postive (or right) skewed than speed.

```{r}
summary(cars)
```

#### Question 1e: What kind of variables are in the cars dataset (i.e. are they nominal, ordinal, ratio, or interval)? (1 pt)

These are ratio variables as they have a true zero.

#### Question 1f: What kind of correlation coefficient would you calculate? Why? (2 pts)

I would calculate the Pearson correlation coefficient because the variables are ratio, continuous variables. If they were categorical, I would have considered Spearman's rank. However, if they are not bivariate normal and contain many outliers, then a Spearman's rank might be more appropriate. A Pearson correlation requires the association to be linear, Spearman's rank doens't. So if there are values that differ substantially from this linear relationship than Spearman's rank is more appropirate.

#### Question 1g: Calculate the appropriate correlation coefficient and interpret it. (2 pts)

The correlation coefficient is 0.8068949. This means there is a __strong__ and __positive__ correlation between speed and dist. This implies that as we increase driving speed, we increase the stopping distance.

```{r}
cor(cars)
```

### Question 2 - Visualize your data

#### Question 2a: Plot one of the variables (i.e. create a marginal plot) and describe the distribution. For the description of the distribution think about whether it looks symmetric, skewed, kurtotic, or normal. (5 pts)

##### For `speed`
The distributiion looks roughly symmetric. There doesn't appear to be much kurtosis and, in general, the distribution looks roughly normal. The plot does indicates that there might be some slightly negative (or left) skew.

```{r}
hist(cars$speed)
```

##### For `dist`
The distribution is definitely positively (or right)-skewed. It does not look normally distributed and it doesn't look kurtotic (i.e. no kurtosis) either. It is not symmetric.

```{r}
hist(cars$dist)
```

#### Question 2b: Create a scatterplot of two variables in the cars dataset and describe their relationship. For the description, is the relationship linear, are their outliers, is there a lot of variability around the line, does this reflect the value of the correlation coefficient you could? (5 pts) 

The relationship between `dist` and `speed` looks roughly linear. In other words, a positive increase in speed corresponds to a roughly positive, linear increase in breaking distance. There is not a lot of variability around the line and the strong, positive linear relationship between the two variables agrees with the strength and direction of the correlation coefficient. Given that the association appears linear between the variables, this also corroborates the use of Pearson's correlation coefficient.

There are a couple of observations that might be considered outliers and there is less observations at the extremes (i.e. at low and high speeds).

```{r}
plot(dist ~ speed, data = cars)
```

### Question 3 - Simple linear regression

#### Question 3a: Run a simple linear regression model. State which variable you used as the response and which you used as the predictor and explain why you made this decision? (2 pt)

Breaking distance is the reponse variable and speed of the car is the predictor. It doesn't make sense to try and predict a car's speed given it's breaking distance but it does make sense the other way around. 

```{r}
mod1 <- lm(dist ~ speed, cars)
```


#### Question 3b: What is the p-value for the predictor (i.e. the slope term)? Please provide an interpretation of this p-value. Did you reject or fail to reject your null hypothesis? (3 pts)

The p-value is 1.49e-12 (or less than .001). The p-value is the probability of obtaining results as extreme or more extreme than those observed given that the null hypothesis is true. So in this case, it is extremely unlikely that the slope of the predictor is 0 given our results. We reject the null hypothesis.

```{r}
summary(mod1)
```

#### Question 3c: Interpret the slope term. (1pt)

The slope is 3.9324. This means that for a one mile per hour increase, the expected increase in breaking distance is 3.9324 feet.

#### Question 3d: Does the 95% confidence interval for your predictor agree with what you found in question 3b (i.e. the p-value and your hypothesis test)? Why? (2 pts)

```{r}
confint(mod1)
```

Yes, the 95% confidence interval for the slope (3.096964, 4.767853) agrees with the findings from the hypothesis test. The reason is that the hypothesis is testing whether 0 is a likely value for the slope. We reject the null hypotheiss and that implies that 0 is an extremely unlikely value based on our p-value. Our 95% confidence interval tells us a range of plausible values for the true slope. 0 is not in the range of plausible values because it isn't between 3.096964 and 4.767853. Therefore, this agrees with the hypothesis test.  

#### Question 3e: Plot your residuals; describe your residual plot; and comment on whether your model is appropriate or not based on this plot. (3 pts) 

In general, the residual plot looks OK. However, there does appear to be increase in variability as you move from the left to the right. In other words, it looks like the residuals are fanning out. This could just be a function of the small number of observations when the speed of the car is slow, so this residual plot is probably OK and that would mean that our model is OK. However, it might be worth it to consider some transformations of our response variable and see what effect this might have on our findings. 


```{r}
plot(mod1, which = 1)
```
