---
title: 'R Assignment #2'
author: "Your Name"
date: "17 September 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

### Brief introduction to R Markdown 

We will be using R Markdown for the R assignments. Why? Because it is an easy way to keep your answers and your R code in one place and it is easier for me to understand what you did and if you went astray. It will make help you make connections between your answers and your R code more seamless.

You can find out lots of stuff about R Markdown online and you can use this cheat sheet available at <https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>. 

For this class, you only need to know two things. 

First, the hashtags correspond to headings. For example,

# First heading
## Second heading
### Third heading

and so up to six heading. You only need to use one level of heading to delineate questions as I've done below. But if you want to better organize your homework you are welcome to do so.

The second thing you need to know is that to include R code you need to create a special environment by doing the following:

```{r}
# ```{r} - this initiates the special environment 
# Example of how you include R code and add 2 + 2 together
2 + 2
# ``` - this terminates the special environment
```

You can insert this by clicking "Chunks" -> "Insert Chunk"

Try not to get to confused by the hashtag in the R code. That is a comment and not a heading. It is there just for me to explain what the code is doing. 


The final thing you need to know is that you need to press "Knit HTML" to actually create your document. You will send your homework to me as an HTML file.

That's everything you need to know about R Markdown. 

### Note about running R code
Always, always write R code in this window! Never write it in the window below. When you type R code in the special environment, you can run a line of code without needing to hit "Knit HTML". You can do this in two ways:

1. Hit the "Run" button
2. Hit Ctrl+Enter on a Windows computer or Command+Enter on a Mac

This is very good for quickly testing that your code is working!

You can safely delete everything in the Introduction heading (including the Introduction header) through the R Assignment 1 header below before you turn in R assignment 1. In fact, that would be nice :) 

# R Assignment 2

For this assignment, you'll be calculating and interpreting various reliability and validity estimates within the classical test theory framework. As with before, we'll use an example to create all the `R` code you'll need for this assignment. The data set used in this example is called `score` from the `irtoys` package and I have saved it in the `score.csv` data set on Piazza. __Please download this data set before starting the example__.

Here is how you read in the dataset:

1. Click the "Environment" tab in the upper right pane of Rstudio
2. Click the "Import Dataset"
3. Click "From Text File..."
4. Locate the `scores.csv` file. This is likely just in Downloads or where ever you put it.
5. Make sure "Heading" is set to "Yes"
6. Click "Import"

To verify that this worked type: `head(score)`. If you don't get an error then everything is OK.

##### Note: If you are on a Windows computer and you are getting an error message or you want to run this using code and not via menus. Run the following:
```{r}
# Reads in the score data set
score <- read.csv("http://cddesja.github.io/classes/e411prma2015-3/data/score.csv", header = T)

# Reads in the Harry Potter data set
hp <- read.csv("http://cddesja.github.io/classes/e411prma2015-3/data/hp.csv", header = T)
```

#####YOU MUST RUN THIS COMMAND BEFORE YOU START THIS ASSIGNMENT! 
```{r}
source("http://cddesja.github.io/classes/e411prma2015-3/data/reliability.R")
```
So, __highlight that command and click "Run" before proceeding!__

```{r, echo = F, eval = F}
# The special arguments above are just to tell R Markdown not to print this in your final HTML document
#

# To calculate the total test scores
total_test <- rowSums(score)

# To plot the total test scores
hist(total_test)

# To find the correlation between the total test scores and the items
cor(score, total_test)

# To get summary information about the dataset
summary(score)

# To get it for just a variable
summary(score$Item1)

# Get a correlation matrix. Note that correlation of a variable with itself is 1
cor(score, use = "complete.obs")

# To split the data set into two
first_split <- score[,seq(1, ncol(score), 2)]
second_split <- score[,seq(2, ncol(score), 2)]

# To create total test scores
first_split$total <- rowSums(first_split)
second_split$total <- rowSums(second_split)

# To plot a variable:
hist(first_split$total)

# To find the correlation between the total test scores
cor(first_split$total, second_split$total)

# To calculate coefficient alpha
coef_alpha(score)

# To calculate KR20
kr20(score)

# If you want to know how long a test should be given a desired reliability of 0.8, a current reliability of 0.75, and a current test length of 40.
new_length(desired = 0.8, present = 0.75, test_length = 40)
# Do note, that you should always round up!

# To find sem first calculate variance of total test scores
sigma <- var(total_test)

# Then use a reliability estimate, can use any reliabilty estimate, but we'll using coef_alpha here.
rel <- coef_alpha(score)
sem(sigma, rel)
```

For this assignment, we'll be using the Harry Potter pretest. Make sure you have downloaded the test, the Rmd file, and the `hp.csv` file. Note, that after you've read in the data, the data will be stored as `hp` unless you've changed something!

Please answer your questions below the question headings and include all the R code that you used to answer the question (if R code was required) below your response.

### Question 1 - Describing the test

#### Question 1a: Did everyone answer every question on the test? (1 pt)

__After you are done answering this question run the following command inside a Chunk __  

`
hp <- na.omit(hp)
`

#### Question 1b: Which item was the easiest and which item was the hardest on the test? (2 pts)

#### Question 1c: What is strange about item 4? Provide an explanation of what could be wrong with this item. Hint: take a look at the test question. (2 pts) 

#### Question 1d: Choose one item from the test and describe it. Describe it's mean, it's variance, and it's distribution by plotting the variable. (3 pts)

#### Question 1e: Which two items are the most correlated? Which two variables are the least correlated? Conceptually, do these correlations make sense to you? Hint: Again take a look at the test questions. (3 pts)

#### Question 1f: Given what you know about the test questions (and if you didn't take the test, please read the questions now), do you think the test is unidimensional? Why or why not? (2 pts) _This question does not require R_.

### Question 2 - Total Test Scores

#### Question 2a: What is the range of the total test scores? Did any of your classmates get all the items correct or all the items incorrect? (2 pts) 

#### Question 2b: Describe the distribution of the test scores. Make sure to talk about skewness, shape, and kurtosis. Please include a plot of this distribution. (3 pts)

#### Question 2c: What item is most highly correlated with the total test score? Provide an interpretation of this correlation. (2 pts)

#### Question 2d: Why do you think item 7 says "NA"? (1 pt)

### Question 3 - Reliability

#### Question 3a: Calculate split-half reliability (1 pt).

#### Question 3b: Provide an estimate of internal consistency and interpret this estimate. (2 pt) 

#### Question 3c: How long should the test be if we want to have a reliability estimate of 0.8 (1pt)?

#### Question 3d: Person 30 had a total test score of 11. Calculate a 95% confidence interval for their true score and provide an interpretation of this confidence interval. (3 pts)

#### Question 3e: What would be a plausible true score for Person 30 and what would be an implausible score for Person 30? (2 pts)

#### Question 3f: Person 13 had a total test score of 5. Did Person 30 do significantly better than Person 13? (3 pts) 

#### Question 3g: Provide two recommendations of how you would improve this test based on your analysis?  (2 pts)