---
title: "Introduction to R"
author: "Christopher David Desjardins"
date: "August 20, 2015"
output: html_document
---
# Section 1. Customizing Rstudio

Much of the Rstudio IDE is customizable. By default Rstudio has soft-wrap R sources files disabled. This might be something you wish to enable to prevent very long lines. I would recommend just breaking your lines after 80 characters or so anyways. 

You can find these settings under Tools -> Global Options.

You can also change the pane layout if you want.

By default, the following layout is used.
Upper Left - Code editor
Bottom Left - R console 
Upper Right - Environment, i.e. what is currently in R and History, i.e. what commands you have run.
Bottom Right - Viewer for help pages, files in current directory, graphs, and packages. 

You'll always want write code in the Code editor and not directly in the R console unless it's something you know you'll never want to write again!

# Section 2. R Basics

First, here are some books.
- Using R for Introductory Statistics (free): http://cran.r-project.org/doc/contrib/Verzani-SimpleR.pdf
- Linear models with R: http://www.amazon.com/Linear-Models-Chapman-Statistical-Science/dp/1584884258
- Extending the linear model with R: http://www.amazon.com/Extending-Linear-Model-Generalized-Nonparametric/dp/158488424X

However, there's lot of R resources. So unless you want a printed R book you don't need to buy one. What I do is often just Google my problem and I either hitstackoverflow or the R-help mailing list. There are also lots of wonderful blogs. 

For example, http://www.statmethods.net/

For Icelandic R videos: https://www.youtube.com/channel/UClpgj8pTFS3XT5SHYWpAbrQ. These are R videos and statistics tutorials for the stats department at HÍ.

R is both a programming language and a statistics environment

To install a library in R use the install.package() function and R will figure out all it's dependencies automatically.

```{r, eval = FALSE}
install.packages("irtoys")  # IRT models
install.packages("lme4")  # HLM aka mixed effects modeling, can do Rasch modeling
install.packages("lavaan")  # CFA, SEM simple output/model syntax to MPlus
install.packages("ggplot2")  # Graphing package 
install.packages("psych")  # CTT, FA, http://personality-project.org/r/
install.packages("plyr")  # Data manipulation
install.packages("equate")  # Equating
```

To load a library.
```{r}
library("foreign")  # This is library needs to be loaded to read SPSS data
```

To list the functions in a particular package
```{r}
lsf.str("package:foreign")
```

More detailed handouts, called vignettes, are available for packages.

```{r, eval = FALSE}
vignette()
vignette("release")  ## ggplot2 vignette
```

To find out information about a function:
```{r, eval = FALSE}
?read.spss  # In spite of the cavaet, I have had no issues with this function
```

To search for a keyword
```{r}
??item
```

What data sets are available?
```{r, eval = FALSE}
data()
```

Which data sets belong to irtoys?
```{r}
data(package = c("irtoys"))
```

To check what directory we're working from and set it
```{r, eval = FALSE}
getwd()
setwd("~/Dropbox/cdd/GitHub/e411prma/data/")
```

To see what's in that directory
```{r}
dir()
```

Import the Harry Potter pretest 

```{r}
# "<-" this is the assignment operator
pretest <- read.csv(file = "~/Dropbox/cdd/GitHub/e411prma/data/hp_pretest.csv", na.strings = "")
# I recomend always naming objects in lowercase.
# If you want to combine words use lower.case
```

Note that pretest appears in the Environment window (UR). You can also see that by running ls() in R
```{r}
ls()
```
You can also note the History tab in the UR that includes all the commands you've run.

Investigate the first few cases and then the last few case.

```{r}
head(pretest)
tail(pretest)
```

To view all the data type
```{r}
pretest
```

If you want to see a specific observation, say #20, and you want to see responses from questions 2 through 3.
```{r}
pretest[20,2:3]
```

Or if you want to see questions 1, 3, and 6 for observations 19 through 21.
```{r}
pretest[19:21, c(1,3,6)]
```

To get information about the data set and the variable types
```{r}
str(pretest)
```
Notice, that these are being treated as factors.

To get the name of all the variables
```{r}
names(pretest)
```

To save the data to a text file named _pretest.txt_ in your working directory
```{r, eval = FALSE}
write.table(pretest, file = "pretest.txt")
```

and to read the file back in
```{r, eval = FALSE}
read.table(file="pretest.txt")
```

We should always read data into an object
```{r, eval = FALSE}
pretest.tmp <- read.table(file="pretest.txt")
```

Similar functions exist to read data directly from CSV files, SAS files, etc. Check out the _foreign_ package. 

To get summary of the pretest data
```{r}
summary(pretest)
```

To access a variable in the data.frame use the "$" symbol. 
```{r}
pretest$Q1  # fyi, R is case sensitive
```

To create a table of this.
```{r}
table(pretest$Q1)
```

Let's a unique ID variable to this data set
```{r}
pretest$id <- 1:nrow(pretest)  # nrow(), number of rows in the data.frame
head(pretest)
```

What if we found out subject 1 should had a 'b' Q3 not 'c'
```{r}
pretest[pretest$id == 1,]
pretest[pretest$id == 1, "Q3"]
pretest[pretest$id == 1, "Q3"] <- "b"
pretest[pretest$id == 1, "Q3"]
```
  
We can use the psych library to get even more information 
```{r}
library("psych")
?describe
describe(pretest)  # asterisks indicate factors
```
But is the appropriate for this kind of data?

Conditional tables, these are especially useful for factors
```{r}
table(pretest$Q1,pretest$Q24)
```

How many levels of factor are there and lists the levels
```{r}
nlevels(pretest$Q2)
levels(pretest$Q2)  
```

Completely remove a variable
```{r}
pretest$Q15 <- NULL
```

# Section 3. Graphing

Graphing can be accomplishing using just the R base libraries. But, graphing is greatly expanded with ggplot2. I'll just show you with ggplot2.

## Scatterplots

The data I will use comes from the mpg dataset. This has miles per gallon. 

```{r}
library(ggplot2)
qplot(x = cty, y = hwy, data = mpg)
```

That command is identical to the previous one.
```{r}
ggplot(aes(x = cty, y = hwy), data = mpg) + geom_point()
```

The advantage of the second line is that you'll often build up and change graphs.
```{r, eval = FALSE}
g0 <- ggplot(aes(x = cty, y = hwy), data = mpg)
g0 + geom_point() + theme_bw()  # let's give it a black and white theme
g0 + geom_point(col = "orange") + theme_bw()  # bw theme and orange points
(g1 <- g0 + geom_point(col = "orange", shape = 2) +
  theme_bw())  # bw theme and orange triangles, () echo
(g1 <- g0 + geom_point(aes(colour = manufacturer), shape = 2) +
   theme_bw()) 
```
Add a best fit
```{r, eval = FALSE}
(g2 <- g1 + stat_smooth(aes(color = manufacturer),method = "lm", formula = y ~ x))
```

Continue to modify the graph and polish it a little bit
```{r, eval = FALSE}
# Continue to modify the graph and polishing it a little
g2 + xlab("city miles per gallon") + ylab("highway miles per gallon") + 
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  ggtitle("Highway Mileage vs. City Mileage")
```

This is how to make box plots.
```{r, eval = FALSE}
ggplot(aes(y = hwy, x = manufacturer), data = mpg) + geom_boxplot()
ggplot(aes(y = hwy, x = manufacturer), data = mpg) +
  geom_boxplot(aes(fill = manufacturer), outlier.color = "green" )
```

Here are some marginal plot of highway mileage
```{r, eval = FALSE}
ggplot(data = mpg, aes(hwy)) + geom_histogram(fill="white", color = "red") +
  xlab("Highway Mileage") + ylab("Frequency")

ggplot(data = mpg, aes(hwy)) + geom_histogram(fill="white", color = "red") +
  xlab("Highway Mileage") + ylab("Frequency") + facet_grid(.~class)

ggplot(data = mpg, aes(hwy)) + geom_histogram(aes(fill=class, color = class)) +
  xlab("Highway Mileage") + ylab("Frequency") + facet_grid(manufacturer~class) 
```

Extra plotting help visit: http://docs.ggplot2.org/current/. Lots of great information is available there.  